---
title: "R Assignment - Estimation and CIs"
author: "Sanchit Vijay"
# date: "today"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

This exercise uses the graduate school admissions dataset from the applicants to a higher-ed institute. The dataset is *LogRegAdmit.csv.*

The variables in the dataset are:  

* `admit`: 0 or 1 (False or True)
* `gre`: gre score
* `gpa`: grade point average at undergraduate level
* `rank`: the ranking of the applicant's undergraduate institute, 1, 2, 3, or 4


## Grad School Admissions

### Question 1  

**After importing the dataset, check the structure. Make sure `admit` and `rank` are stored as factors, instead of numeric or integers. (Hint: Use as.factor() function or factor() function).**
```{r, results='markup'}
admissions = data.frame(read.csv("LogRegAdmit.csv"))
admissions$admit = as.factor(admissions$admit)
admissions$rank = as.factor(admissions$rank)
head(admissions)
cat("Datatype of admit: ", class(admissions$admit))
cat("\n")
cat("Datatype of rank: ", class(admissions$rank))
```

'admit' and 'rank' attributes are stored as factors.  

### Question 2  
**Construct t-intervals for the `gre` and `gpa` variables for all applicants at 0.80 level and 0.99 level.**
```{r, results='markup'}
t.test(x = admissions$gre, conf.level = 0.8)
t.test(x = admissions$gpa, conf.level = 0.8)

t.test(x = admissions$gre, conf.level = 0.99)
t.test(x = admissions$gpa, conf.level = 0.99)
```

### Question 3  
**Repeat the same calculation for `gre` in Question 2 but for admitted (1) and rejected (0) separately. (You can subset them first.)  Between the admitted and rejected, do the two intervals overlap at 0.80 level? And at 0.99 level?**
```{r, results='markup'}
gre_subset_adm = subset(admissions, admit == 1)
t.test(x = gre_subset_adm$gre, conf.level = 0.80)
t.test(x = gre_subset_adm$gre, conf.level = 0.99)

gre_subset_rej = subset(admissions, admit == 0)
t.test(x = gre_subset_rej$gre, conf.level = 0.80)
t.test(x = gre_subset_rej$gre, conf.level = 0.99)
```
Intervals don't overlap between admitted and rejected region for both 0.80 and 0.99 confidence levels.  
  
### Question 4   
**Make (box-) plots showing `gre` distribution among applicants from different school rankings for the admitted and rejected separately. Please use ggplot for this. The x-variable should be `rank`, and the y-variable should be `gre`.**
```{r, results='asis'}
library(ggplot2)

ggplot(gre_subset_adm, aes(x = rank, y = gre)) + geom_point(color = "blue") + labs(title = "GRE score vs. school rank distribution of admitted applicants", x = "School ranking", y = "GRE score")

ggplot(gre_subset_rej, aes(x = rank, y = gre)) + geom_point(color = "pink") + labs(title = "GRE score vs. school rank distribution of rejected applicants", x = "School ranking", y = "GRE score")
```

### Question 5   
**Repeat Question 2 and Question 3 for `gpa`. Do the two groups have t-intervals overlap at 0.80 level? At 0.99 level?**
```{r, results='markup'}
gpa_subset_adm = subset(admissions, admit == 1)
t.test(x = gpa_subset_adm$gpa, conf.level = 0.80)
t.test(x = gpa_subset_adm$gpa, conf.level = 0.99)

gpa_subset_rej = subset(admissions, admit == 0)
t.test(x = gpa_subset_rej$gpa, conf.level = 0.80)
t.test(x = gpa_subset_rej$gpa, conf.level = 0.99)
```
Intervals don't overlap between admitted and rejected region for both 0.80 and 0.99 confidence levels.  

### Question 6   
**Repeat Question 4 for `gpa`. Make (box-) plots showing the gpa distribution among applicants from different school rankings for the admitted and rejected separately. Again, use ggplot.**
```{r, results='markup'}
ggplot(gpa_subset_adm, aes(x = rank, y = gpa)) + geom_point(color = "cyan") + labs(title = "GPA vs. school rank distribution of admitted applicants", x = "School ranking", y = "GPA")

ggplot(gpa_subset_rej, aes(x = rank, y = gpa)) + geom_point(color = "green") + labs(title = "GPA vs. school rank distribution of rejected applicants", x = "School ranking", y = "GPA")
```

### Question 7  
**Based on EDA from the charts and the t-interval calculations you produced, do you observe any potential effects between gre/gpa/rank on admissions? Explain briefly. **
```{r, results='markup'}
# ggplot(gpa_subset_adm, aes(x = gre, y = gpa, color = rank)) + geom_point() + labs(title = "GPA vs. school rank distribution of admitted applicants", x = "GRE", y = "GPA")
# ggplot(gpa_subset_rej, aes(x = gre, y = gpa, color = rank)) + geom_point() + labs(title = "GPA vs. school rank distribution of admitted applicants", x = "GRE", y = "GPA")
```

There are few observations made from the above EDA and t-interval calculations. If we look at GRE score and GPA t-intervals for admitted applicants it is higher and non-overlapping for both the confidence levels. Higher GRE score above 594 is required but is not the only factor that affects chances of receiving an admit. Similarly GPA of 3.4 is expected but does not guarantee an admit. Applicants with school rank of 1 with higher GRE score are also rejected maybe because of their low GPA. Conclusively, GRE, GPA, and ranking significantly affects the chances of receiving an admit but are not the only factors. 




